15_networks

### 2019-11-13 analysis/15_networks.Rmd
new plots with manipluation and candidate genes

```{r boxplots}
# genes from website shown to be associate with oxt and prl
OXTgenes <- c("OXT",  "AVPR1A", "AVPR1B", "TRH", 
              "HCRT", "GNRH1", "CRH", "POMC", "MC4R", "AVPR2")
PRLgenes <- c("PRL", "GNRH1", "TRH", "POU1F1", "POMC", "TSHB",
              "FSHB", "GH", "PRLR", "JAK2")
OXTPRLgenes <- c(OXTgenes, PRLgenes)
head(OXTPRLgenes)

@@

longdata <- widecountdata %>% 
            select(V1:study, OXTPRLgenes, PRL_associated) %>%
            gather(gene, expression, OXT:UBE2C)
head(longdata)

plotnetworkgenes <- function(whichgenes){
  longdata %>%
  filter(gene %in% whichgenes) %>%
  ggplot(aes(x = treatment, y = expression)) +
        geom_smooth(aes(colour = sex)) +
    geom_boxplot(aes(outlier.color = treatment, 
                     fill = treatment, alpha = sex)) +
    mytheme() +
    theme(legend.position = "none",
          axis.text.x = element_blank()) +
    scale_fill_manual(values = colorschar) +
    scale_color_manual(values = sexcolors) +
    labs(x = NULL) +
    scale_alpha_manual(values = c(0.75,1)) +
  facet_wrap(~gene, scales = "free_y")
}

plotnetworkgenes(OXTgenes)
plotnetworkgenes(PRLgenes)
plotnetworkgenes(PRL_associated[85:93])

@@

```{r correlations}
forcorrdata <- widecountdata %>% 
            select(V1, OXTPRLgenes)  %>% 
            as.data.frame()
row.names(forcorrdata) <- forcorrdata$V1   
forcorrdata$V1 <- NULL
head(forcorrdata)

@@

```{r wgcnaPRLgenes}
x <- widecountdata %>% 
            select(PRL_associated,OXTPRLgenes) %>%
       correlate()     # Create correlation data frame (cor_df)

>>>>

```{r boxplots}
# genes from website shown to be associate with oxt and prl
PRLgenes <- c("PRL", "GNRH1", "TRH", "POU1F1", "POMC", "TSHB",
              "FSHB", "GH", "PRLR", "JAK2", "SOCS1", "CISH")
OXTgenes <- c("OXT",  "AVPR1A", "AVPR1B", "TRH", 
              "HCRT", "GNRH1", "CRH", "POMC", "MC4R", "AVPR2")
CISHgenes <- c("CISH", "SOCS1", "SOCS2", "SOCS4", "SOCS5", "SOCS6")

PRLnOther <- c(PRLgenes,OXTgenes, CISHgenes)

@@

# cancer genes from PRL list from  http://bioinformatics.sdstate.edu/go/
# shiny go for Gene Ontology Enrichment
cancergenes <- c("E2F1", "MYC" ,"PLK1", "BRCA1" ,"AURKA", "PRL")
cellcyclegenes <- c("E2F1", "MAD2L1", "MCM2", "PLK1", "MCM10", "BUB1", 
                    "CDC45", "CCNB3", "CCNA2", 
                      "CDK1", "CDC6", "CDC20")
replicationanddamage <- c("MCM2"  , "CDT1" , " CDC45 " , "CDC6", 
                          "E2F1" , " MYC" , "BRCA1" , "CCNB3 " , "CDK1")

## plot genes

longdata <- widecountdata %>% 
            select(V1:study, PRLnOther, PRL_associated) %>%
            gather(gene, expression, PRL:UBE2C)
head(longdata)

plotnetworkgenes <- function(whichgenes, mysubtitle){
  longdata %>%
  filter(gene %in% whichgenes) %>%
  ggplot(aes(x = as.numeric(treatment), y = expression)) +
        geom_smooth(aes(colour = sex)) +
    geom_boxplot(aes(fill = treatment, alpha = sex)) +
    mytheme() +
    theme(legend.position = "none",
          axis.text.x = element_blank()) +
    scale_fill_manual(values = colorschar) +
    scale_color_manual(values = sexcolors) +
    labs(x = NULL, subtitle = mysubtitle) +
    scale_alpha_manual(values = c(0.75,1)) +
  facet_wrap(~gene, scales = "free_y")
}

plotnetworkgenes(PRLgenes, "Known PRL associates") 
plotnetworkgenes(PRL_associated[85:93],  "in WGCNA module with PRL")
plotnetworkgenes(cancergenes, "genes in Breast Cancer WikiPathway")
plotnetworkgenes(cellcyclegenes, "genes in Cell Cycle WikiPathways ")

plotnetworkgenes(replicationanddamage, 
                 "genes in DNA Replication & Damage Response WikiPathway")

@@

```{r correlations}
forcorrdata <- widecountdata %>% 
            select(cancergenes)  %>% 
            as.data.frame()
row.names(forcorrdata) <- forcorrdata$V1   
forcorrdata$V1 <- NULL
head(forcorrdata)

@@

```{r wgcnaPRLgenes}
x <- widecountdata %>% 
            select(PRL_associated,PRLnOther) %>%
       correlate()     # Create correlation data frame (cor_df)
